---
title: "Case Study 01 WC + JT"
author: "Joe Chung w/Jonathan Tan"
date: "6/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#setwd("D:/SMU/DS 6306 Doing Data Science/DS_6306_case_study_01/DS_6306_case_study_01")
```
```{r package installations, include = FALSE}
#install.packages("maps")
#library(ggplot2)
#library(plotly)
#library(dplyr)
#library(maps)
#library(ggthemes)
```
#Introduction
This document is an analysis of the classification and origin of 2410 brands of beer produced by 558 breweries across the continential United States. 

Of particular interest was the high concentration of unique beer brand production in breweries on the west coast, the significantly above average number of IPA style beers, and majority of beers that cluster around low alcohol content and bitterness. 

Colorado has 39 breweries reported in the data, followed by California with 37. Additional information on the destinations of beer produced in these states and further research could determine if this indicates a reigonally well established demand for local beer variety, a thriving interstate export business, or an oversaturated production environment. 

Of the 2410 brands represented in the data, 17.6% of them are IPA style.  The next 10% are APA, and 5.5% American Amber/Pale Ale. The top 3 most populous styles of beer make up more than 1/3 of all the reported beer brands. Further research could determine if this stylistic majority represents a marketplace glut or genuine demand for beers in these styles. 

(percentage from graph) of the beer brands in the data have an alcohol content of 5% or less, with a bitterness rating of 25, with only (percentage) of brands exceeding 10% ABV and 100 IBU. This indicates that most breweries believe that beers around this level of bitterness and alcohol have the highest rate of consumption. Further research could determine if the alcohol content and bitterness level of a beer is a factor in popularity, or if production cost or legal stipulations make production more attractive to breweries.


```{r main, echo=TRUE}
#'x' is the column of a data.frame that holds 2 digit state codes
stateFromLower <-function(x) {
  #read 52 state codes into local variable [includes DC (Washington D.C. and PR (Puerto Rico)]
  st.codes<-data.frame(
    state=as.factor(c("AK", "AL", "AR", "AZ", "CA", "CO", "CT", "DC", "DE", "FL", "GA",
                      "HI", "IA", "ID", "IL", "IN", "KS", "KY", "LA", "MA", "MD", "ME",
                      "MI", "MN", "MO", "MS",  "MT", "NC", "ND", "NE", "NH", "NJ", "NM",
                      "NV", "NY", "OH", "OK", "OR", "PA", "PR", "RI", "SC", "SD", "TN",
                      "TX", "UT", "VA", "VT", "WA", "WI", "WV", "WY")),
    full=as.factor(c("alaska","alabama","arkansas","arizona","california","colorado",
                     "connecticut","district of columbia","delaware","florida","georgia",
                     "hawaii","iowa","idaho","illinois","indiana","kansas","kentucky",
                     "louisiana","massachusetts","maryland","maine","michigan","minnesota",
                     "missouri","mississippi","montana","north carolina","north dakota",
                     "nebraska","new hampshire","new jersey","new mexico","nevada",
                     "new york","ohio","oklahoma","oregon","pennsylvania","puerto rico",
                     "rhode island","south carolina","south dakota","tennessee","texas",
                     "utah","virginia","vermont","washington","wisconsin",
                     "west virginia","wyoming"))
  )
  #create an nx1 data.frame of state codes from source column
  st.x<-data.frame(state=x)
  #match source codes with codes from 'st.codes' local variable and use to return the full state name
  refac.x<-st.codes$full[match(st.x$state,st.codes$state)]
  #return the full state names in the same order in which they appeared in the original source
  return(refac.x)
  
}
```

#1.   How many breweries are present in each state?
```{r q1}
#beer_path="/Users/wailunchung/Documents/GitHub/MSDS-6306-Doing-Data-Science/Unit 7/Beers.csv"
#brewery_path="/Users/wailunchung/Documents/GitHub/MSDS-6306-Doing-Data-Science/Unit 7/Breweries.csv"
#added some redirects for my machine, switch comments back when you run it on yours jtan
beer_path="D:/SMU/DS 6306 Doing Data Science/DS_6306_case_study_01/DS_6306_case_study_01/Raw Data/Beers.csv"
brewery_path="D:/SMU/DS 6306 Doing Data Science/DS_6306_case_study_01/DS_6306_case_study_01/Raw Data/Breweries.csv"
beer_df=read.csv(beer_path)
brewery_df=read.csv(brewery_path)

# returns string w/o leading whitespace
trim <- function (x)  sub("^\\s+", "", x)
brewery_df$State <- trim(brewery_df$State) 

# covert state to full state name
brewery_df$State_full<-stateFromLower(brewery_df$State)
count_by_state <- data.frame(table(brewery_df$State_full))

#ggplot(count_by_state, aes(x=reorder(Var1,-Freq),y=Freq,fill=Var1)) + 
#  geom_bar(stat="identity") + 
#  labs(title="Brewery Count by State",subtitle="",x="State",y="freq",caption="") +
#  theme(plot.title=element_text(hjust=0.5),axis.text.x = element_text(angle = 90, hjust = #1),legend.position="none")

#install.packages("maps")
# using maps to get coordinate and use ggplot to plot maps
#install.packages('ggthemes')
library(maps)
library(ggplot2)
library(ggthemes)
us_states <- map_data("state")
head(us_states,10)
beer_brew_df<-merge(x=beer_df,y=brewery_df,by.x="Brewery_id",by.y="Brew_ID",all.x = TRUE)
#beer_brew_df_map<-merge(x=brewery_df,y=us_states,by.x="State_full",by.y="region",all.x = TRUE)
beer_brew_df_map<-merge(x=count_by_state,y=us_states,by.x="Var1",by.y="region",all.x = TRUE)
#count_by_state
#names(beer_brew_df_map)

#beer_brew_df_map

p <- ggplot(data = beer_brew_df_map,
            aes(x = long, y = lat,
                group = group, fill = Freq))
#p + geom_polygon(color = "gray90", size = 0.1) + guides(fill = FALSE)
p + geom_polygon(color = "gray90", size = 0.1) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) + 
  labs(title = "Brewery Distribution") +
  theme_map()
```


#2. Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.
```{r q2}
beer_brew_df<-merge(x=beer_df,y=brewery_df,by.x="Brewery_id",by.y="Brew_ID",all.x = TRUE)
head(beer_brew_df,6)
tail(beer_brew_df,6)
```

#3. Report the number of NA's in each column.
```{r q3}
summary(beer_df$ABV) #62 na
summary(beer_df$IBU) #1005 na
sapply(beer_brew_df, function(x) sum(is.na(x))) #another check
#optional remove NA's
cmerge <- beer_brew_df[complete.cases(beer_brew_df),]
#sapply(cmerge, function(x) sum(is.na(x))) #to check cmerge for NA's

#filter by abv/ibu to get boxes for slide 8
#less than 100 IBU, 10% ABV
bbox<- subset(cmerge, cmerge$ABV <0.1 & cmerge$IBU < 100)
#1353/1405, 96.2%
#less than 50 IBU, 7.5% ABV
rbox <- subset(cmerge, cmerge$ABV < 0.06 &cmerge$ABV > 0.03 & cmerge$IBU <50)
#740/1405, 52.6%
```
There are 62 beers with no reported ABV, and 1005 beers with no reported IBU

#4. Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.

```{r q4}
#library(dplyr)
#beer_brew_df %>%
#  group_by(State)%>% 
#  summarise(Mean=mean(ABV,na.rm=TRUE), Max=max(ABV,na.rm=TRUE), Min=min(ABV,na.rm=TRUE), Median=median(ABV,na.rm=TRUE), Std=sd(ABV,na.rm=TRUE))

library(dplyr)
median_IBU_by_state <- beer_brew_df %>%
  group_by(State_full)%>% 
  summarise(Median=median(IBU,na.rm=TRUE))

median_IBU_by_state[is.na(median_IBU_by_state$Median),]
# *** state "SD" no Median, All IBU are NA in SD
distinct(beer_brew_df,State)
beer_brew_df[beer_brew_df$State=='SD',]

ggplot(median_IBU_by_state, aes(x=reorder(State_full,-Median),y=Median,fill=State_full)) + 
  geom_bar(stat="identity") + 
  labs(title="IBU Median by State",subtitle="",x="State",y="Median",caption="") +
  theme(plot.title=element_text(hjust=0.5),axis.text.x = element_text(angle = 90, hjust = 1),legend.position="none")
```
National Average ABV is 5.9%
National Average IBU is 42.7

#5. Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r q5}
ABV_by_state <- beer_brew_df %>%
  group_by(State_full)%>% 
  summarise(Mean=mean(ABV,na.rm=TRUE),Median=median(ABV,na.rm=TRUE))

ABV_by_state

IBU_by_state <- beer_brew_df %>%
  group_by(State_full)%>% 
  summarise(Mean=mean(IBU,na.rm=TRUE),Median=median(IBU,na.rm=TRUE))

IBU_by_state

#Top 10 ABV state (Mean)
top_n(ABV_by_state, n=10, Mean) %>%
  ggplot(., aes(x=reorder(State_full,-Mean),y=Mean,fill=State_full)) + 
  geom_bar(stat="identity") + 
  labs(title="Top 10 ABV Average by State",subtitle="",x="State",y="Average",caption="") +
  theme(plot.title=element_text(hjust=0.5),axis.text.x = element_text(angle = 90, hjust = 1),legend.position="none")

#Top 10 IBU state (Mean)
top_n(IBU_by_state, n=10, Mean) %>%
  ggplot(., aes(x=reorder(State_full,-Mean),y=Mean,fill=State_full)) + 
  geom_bar(stat="identity") + 
  labs(title="Top 10 IBU Average by State",subtitle="",x="State",y="Average",caption="") +
  theme(plot.title=element_text(hjust=0.5),axis.text.x = element_text(angle = 90, hjust = 1),legend.position="none")



#--------

summary(beer_brew_df)
median_ABV_by_state <- beer_brew_df %>%
  group_by(State_full)%>% 
  summarise(Median=median(ABV,na.rm=TRUE))

#Top 10 ABV state
top_n(median_ABV_by_state, n=10, Median) %>%
ggplot(., aes(x=reorder(State_full,-Median),y=Median,fill=State_full)) + 
  geom_bar(stat="identity") + 
  labs(title="Top 10 ABV Median by State",subtitle="",x="State",y="Median",caption="") +
  theme(plot.title=element_text(hjust=0.5),axis.text.x = element_text(angle = 90, hjust = 1),legend.position="none") #+ geom_smooth(method = "lm")

#Top 10 IBU state
top_n(median_IBU_by_state, n=10, Median) %>%
  ggplot(., aes(x=reorder(State_full,-Median),y=Median,fill=State_full)) + 
  geom_bar(stat="identity") + 
  labs(title="Top 10 IBU Median by State",subtitle="",x="State",y="Median",caption="") +
  theme(plot.title=element_text(hjust=0.5),axis.text.x = element_text(angle = 90, hjust = 1),legend.position="none")
```
The highest national ABV is (name of beer),(some percent%) alcohol by volume, produced in (state)
The highest national IBU is (name of beer),(some rating) bitterness rating, produced in (state)

#6. Summary statistics for the ABV variable.
```{r q6}
summary(beer_brew_df$ABV)
```

#7. Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.
```{r q7}
ggplot(beer_brew_df, aes(x=ABV, y=IBU)) + 
  geom_point(shape=18, color="blue")+
  geom_smooth(method=lm,  linetype="dashed",
              color="darkred", fill="blue")
```
#Extra: ABV, IBU by style
```{r extra}
beer_df$Style_cat='Unknown'

#clean special characters 
beer_df$Style <- gsub('Kölsch', 'Kolsch', beer_df$Style)
beer_df$Style <- gsub('Märzen', 'Marzen', beer_df$Style)
beer_df$Style <- gsub('Bière', 'Biere', beer_df$Style)

#IPA
beer_df$Style_cat <- ifelse(grepl("India Pale Ale", ignore.case=TRUE, beer_df$Style), "IPA", beer_df$Style_cat)
beer_df$Style_cat <- ifelse(grepl("IPA", ignore.case=TRUE, beer_df$Style), "IPA", beer_df$Style_cat)
#Porters
beer_df$Style_cat <- ifelse(grepl("porter", ignore.case=TRUE, beer_df$Style), "Porters", beer_df$Style_cat)
#Stouts
beer_df$Style_cat <- ifelse(grepl("stout", ignore.case=TRUE, beer_df$Style), "Stouts", beer_df$Style_cat)

#Strong Ales
strong_ales_key <-c('red ale','wine ale', 'old ale', 'wine', 'strong ale','quadrupel','tripel','Scotch Ale','Wee Heavy')
beer_df$Style_cat <- ifelse(grepl(paste(strong_ales_key,collapse="|"), ignore.case=TRUE, beer_df$Style), "Strong Ales", beer_df$Style_cat)

#Wheat Beers
wheat_beer_key <- c('wheat', 'Witbier', 'Weissbier','Dunkelweizen','Hefeweizen')
beer_df$Style_cat <- ifelse(grepl(paste(wheat_beer_key,collapse="|"), ignore.case=TRUE, beer_df$Style), "Wheat Beers", beer_df$Style_cat)

#Wild or Sour Beers
wild_beer_key <-c('Gose', 'wild ale','oud bruin','Flanders red ale')
beer_df$Style_cat <- ifelse(grepl(paste(wild_beer_key,collapse="|"), ignore.case=TRUE, beer_df$Style), "Wild or Sour Beers", beer_df$Style_cat)

#Hybrid Beers
hybrid_beer_key <- c('california common', 'Garde', 'cream ale','kolsch','Altbier','braggot')
beer_df$Style_cat <- ifelse(grepl(paste(hybrid_beer_key,collapse="|"), ignore.case=TRUE, beer_df$Style), "Hybrid Beers", beer_df$Style_cat)

#Dark Lagers
dark_key <-c('schwarzbier','Marzen','Oktoberfest')
beer_df$Style_cat <- ifelse(grepl(paste(dark_key,collapse="|"), ignore.case=TRUE, beer_df$Style), "Dark Lagers", beer_df$Style_cat)

#Bocks
bock_key<-c('Weizen','bock')
beer_df$Style_cat <- ifelse(grepl(paste(bock_key,collapse="|"), ignore.case=TRUE, beer_df$Style), "Bocks", beer_df$Style_cat)

#Dark Ales
dark_ale_key<-c('black ale', 'dark ale', 'dubbel', 'roggenbier', 'scottish ale', 'winter warmer')
beer_df$Style_cat <- ifelse(grepl(paste(dark_ale_key,collapse="|"), ignore.case=TRUE, beer_df$Style), "Dark Ales", beer_df$Style_cat)

#Brown Ales
brown_key<-c('Brown Ale','Oud Bruin','Mild')
beer_df$Style_cat <- ifelse(grepl(paste(brown_key,collapse="|"), ignore.case=TRUE, beer_df$Style), "Brown Ales", beer_df$Style_cat)

#Specialty Beers
spec_key<-c('Shandy','cider','mead','radler','chile beer','low alcohol beer','black ale','Pumpkin Ale', 'Rye Beer', 'Smoked Beer', 'rauchbier', 'herbed','spiced','Fruit','Vegetable')
beer_df$Style_cat <- ifelse(grepl(paste(spec_key,collapse="|"), ignore.case=TRUE, beer_df$Style), "Specialty Beers", beer_df$Style_cat)

#Pale Ales
pal_key<-c('blonde ale', 'bitter', 'ESB', 'pale ale', 'Abbey Single Ale','saison','grisette')
beer_df$Style_cat <- ifelse(grepl(paste(pal_key,collapse="|"), ignore.case=TRUE, beer_df$Style), "Pale Ales", beer_df$Style_cat)

#Pilsener and Pale Lager
pil_key<-c('pilsener', 'pilsner','Helles','Keller Bier','Zwickel Bier','Malt Liquor','american lager')
beer_df$Style_cat <- ifelse(grepl(paste(pil_key,collapse="|"), ignore.case=TRUE, beer_df$Style), 'Pilsener and Pale Lager', beer_df$Style_cat)

#Lager
lager_key<-c(' lager')
beer_df$Style_cat <- ifelse(grepl(paste(lager_key,collapse="|"), ignore.case=TRUE, beer_df$Style), "Lager", beer_df$Style_cat)

beer_brew_df<-merge(x=beer_df,y=brewery_df,by.x="Brewery_id",by.y="Brew_ID",all.x = TRUE)
ABV_by_style <- beer_brew_df %>%
  group_by(Style_cat)%>% 
  summarise(Mean=mean(ABV,na.rm=TRUE),Median=median(ABV,na.rm=TRUE))

ABV_by_style

IBU_by_style <- beer_brew_df %>%
  group_by(Style_cat)%>% 
  summarise(Mean=mean(IBU,na.rm=TRUE),Median=median(IBU,na.rm=TRUE))

IBU_by_style

#Top 10 ABV Style (Mean)
top_n(ABV_by_style, n=99, Mean) %>%
  ggplot(., aes(x=reorder(Style_cat,-Mean),y=Mean,fill=Style_cat)) + 
  geom_bar(stat="identity") + 
  labs(title="Top 10 ABV Average by Style",subtitle="",x="Style",y="Average",caption="") +
  theme(plot.title=element_text(hjust=0.5),axis.text.x = element_text(angle = 90, hjust = 1),legend.position="none")

#Top 10 IBU Style (Mean)
top_n(IBU_by_style, n=99, Mean) %>%
  ggplot(., aes(x=reorder(Style_cat,-Mean),y=Mean,fill=Style_cat)) + 
  geom_bar(stat="identity") + 
  labs(title="Top 10 IBU Average by Style",subtitle="",x="Style",y="Average",caption="") +
  theme(plot.title=element_text(hjust=0.5),axis.text.x = element_text(angle = 90, hjust = 1),legend.position="none")



```

##extra2 style pie chart
```{r pie}
#install.packages("ggforce")
library(dplyr)
style1 <- beer_df$style

popular.styles <- table(unlist(beerdata$Style))
#want to get this in order, descending
#maybe dataframe can sort and keep label
pop.df <- as.data.frame(popular.styles)
names(pop.df) <- c("Style", "Count")
popular.styles <- pop.df[order(pop.df$Count, decreasing = TRUE),]
#ok got it
pstyles <- popular.styles

#calculate percentage of reported data per style
plabel <- data.frame(pstyles$Style, pstyles$Count/sum(pstyles$Count)*100)
names(plabel) <- c("Style", "percent")

p <- plot_ly(pstyles, labels = pstyles$Style, values = plabel$percent, type = 'pie',textposition = 'outside',textinfo = 'label+percent') %>%
  layout(title = 'Style Distribution',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
#too many one instance styles, trim?

pstyles.trim <- pstyles[1:30,]
plabel.trim <- plabel[1:30, ]
p2 <- plot_ly(pstyles.trim, labels = pstyles.trim$Style, values = plabel.trim$percent, type = 'pie',textposition = 'outside',textinfo = 'label+percent') %>%
  layout(title = 'Style Distribution',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
```
big high level conclusion animated slide at the end, conclusion, 
thank you for the opportunity to present, deeper questions can be reached at email/contact informations.
slide design good
keep to talking points on the left, let visuals speak for themselves
polish and practice, 5 times walkthrough, then video till it has no flaws 
plot a median line on charts where it is a relevent talking point

analysis factors, add percentage of missing data beers